Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-07-17 paper_jasa.tex
color back in fig 10

%Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{eqn:mixed}).   
\noindent The estimated subject-specific power curves are shown in \blue{Figure \ref{fig:power_mixed_subject} (light grey lines). The blue curve is the overall} average power estimated from the model, and for comparison, the grey dashed lines are the power curves for the conventional test. Subject-specific power is quite different for three experiments. In experiment 1 subjects performed similarly, and substantially worse than the conventional test. In experiment 2 there is more variability between subjects, with some doing better than the conventional test on small effects. In experiment 3 there is the most subject-specific variation. Some subjects performed substantially better than the conventional test, and on average the visual test was better.

@@

In the real setting, where visual inference is to be useful, there will be no conventional tests $p$-values. Assessing the strength of perceived structure is a critical component of visual inference. In experiments 1 and 2, there is a $p$-value associated with the actual data plot in each lineup. As the $p$-value increases the proportion of correct responses falls (Figure \ref{fig:pval_pcorrect}), which is evidence of direct association between proportion of correct responses and conventional test $p$-values. When $p$-value is larger than 0.15, it is very uncommon for subjects to correctly identify the actual data plot in the lineup.

@@

From the experimental data the visual $p$-values are computed using Definition \ref{dfn:pvalue}. Figure \ref{fig:pval_definition} displays this visual $p$-value for each lineup against the conventional $p$-value.  The pattern of visual $p$-value is interesting. If there is a signal in the data plot, visual $p$-value has a tendency to take a smaller value (small enough to reject $H_0$) and if not it takes a larger value (large enough to not reject $H_0$) giving a clear indication to reject $H_0$ or not. This is why we do not see lot of visual $p$-values in between 0.05 and 0.8 especially for experiment 2. This guide the researcher to make decision confidently while conventional $p$-values take some marginal $p$-values for which researchers get confused whether to reject or 'not reject' $H_0$. For visual test this is not common.

>>>>

%Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{eqn:mixed}).   
\noindent 
Figure  \ref{fig:power_mixed_subject} shows power curves for both the overall experiment and subject-specific variations. The thick continuous line shows overall estimated power, the thinner lines correspond to subject-specific power curves. For comparison, the  dashed lines show power curves of the conventional test. Subject-specific power is quite different between the three experiments. In experiment 1 subjects performed similarly, and substantially worse than the conventional test. In experiment 2 there is more variability between subjects, with some doing better than the conventional test on small effects. In experiment 3 there is the most subject-specific variation. Some subjects performed substantially better than the conventional test, and on average the visual test was better.

@@

In the real setting, where visual inference is to be useful, there will be no conventional tests $p$-values. Assessing the strength of perceived structure is a critical component of visual inference. In experiments 1 and 2, there is a $p$-value associated with the actual data plot in each lineup. As the $p$-value increases the proportion of correct responses falls (Figure \ref{fig:pval_pcorrect}), which is evidence of direct association between proportion of correct responses and conventional test $p$-values. For  $p$-values  larger than 0.15 it is very uncommon for subjects to correctly identify the actual data plot in the lineup.

@@

From the experimental data the visual $p$-values are estimated based on Definition \ref{dfn:pvalue}. Figure \ref{fig:pval_definition} displays resulting estimates for each lineup against the conventional $p$-value.  The pattern of visual $p$-values is interesting: for small $p$-values the visual estimates tend to be very small, while lineups with larger $p$-values result in very large visual estimate, 
% If there is a signal in the data plot, visual $p$-value has a tendency to take a smaller value (small enough to reject $H_0$) and if not it takes a larger value (large enough to not reject $H_0$) 
% 
 giving a clear indication to reject $H_0$ or not. This is why we do not see lot of visual $p$-values in between 0.05 and 0.8 especially for experiment 2. This guide the researcher to make decision confidently while conventional $p$-values take some marginal $p$-values for which researchers get confused whether to reject or 'not reject' $H_0$. For visual test this is not common.
